### 2021-07-27 text/suppl_methods.Rmd
edits methods from thesis

---
output:
  word_document:
    fig_caption: yes
    reference_docx: ../text/word_template.docx
bibliography: ../text/global_proc.bib
csl: nature.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\usepackage[dvipsnames]{xcolor}

# Supplementary methods

To apply the bioenergetic model that estimates fluxes of carbon (C), nitrogen (N), and phosphorus (P), a number of parameters are required [see @Schiettekatte2020]. Here, we describe how these parameters were quantified for all 1110 species in our database, with a combination of literature, empirical measures, and Bayesian models. All protocols related to the capture and handling of fish complied to the ethical standards of CRIOBE and EPHE, and the University of California Santa Barbara’s Institutional Animal Care and Use Committee (IACUC #915 2016-2019). Extraction and transport of samples were approved by the government of French Polynesia. All analyzes were carried out in R v.3.6.3 and Bayesian modes were run using Stan [@Carpenter2017] and the R package brms [@Burkner2017].

## 1. Growth parameters

The developed bioenergetic model `fishflux` uses the Von Bertalanffy growth function to calculate daily growth.
This growth function relies on three parameters: $l_{\infty}$, $\kappa$, and $t_0$.
For $\kappa$, we used a standardized coefficient that describes the potential growth trajectory of an individual if $l_{\infty}$ were to be equal to its maximum length [i.e., $k_{max}$, @Morais2018].
For simplicity, we kept $t0$ constant at 0 for all species.

### 1.1 Data compilation
We first compiled maximum lengths for all species using Fishbase [@Froese2018] and used these lengths for the $l_{\infty}$. Then, we extracted the available data for $k_{max}$ from Morais & Bellwood (2018), which are essentially $\kappa$ values from Fishbase, standardized to the maximum length [@Morais2018]. We filtered out the species of our species list and only included the $k_{max}$ estimates coming from otolith studies.
In total, this selection process resulted in 439 observations of kmax for varying species and temperatures. \par

Further, we used an openly accessible otolith dataset, including measurements of fishes from five French Polynesian islands, and to which we contributed through data collection and analysis [@Morat2020].
This dataset includes total length measurements to the nearest millimeter and fish weights to the nearest 0.1 grams. Further, it includes distances between annual growth increments, measured by two independent researchers to prevent biases induced by a single observer.
These data can be used to estimate growth rate parameters using a two-step approach: back-calculation to achieve individual-level size-at-age data and hierarchical regression to fit the Von Bertalanffy growth curve (see REF FISHGROWBOT). \par

Specifically, we used the Modified Fry back-calculation model (MF) [@Vigliola2000] to estimate fish lengths at previous ages. We adapted the traditional model to also estimate the uncertainty around the obtained length estimates and allow for the inclusion of missing values using a Bayesian approach (see ref fishgrowbot for details). Then, we fitted the Von Bertalanffy growth model to all species at each location for which there were at least 3 individuals, using Bayesian hierarchical regression models. Both steps of this procedure were carried out using the developed R package 'fishgrowbot' (ref). The package also automatically estimates the parameter $k_{max}$. \par

After combining the two data sources, we obtained 496 estimates of $k_{max}$ for 181 species across varying temperatures.

### 1.2 Data analysis and extrapolation

\noindent Aside from phylogeny, $k_{max}$ is mostly determined by body size and temperature [@Morais2018].

We applied a Bayesian hierarchical model to predict the growth rate of fishes as a function of body size, temperature and phylogeny:

<!-- \begin{equation} -->
$$\textrm{ln}kmax = (\beta_{\textrm{0}} + \gamma_{\textrm{0phy}}) +  \beta_{\textrm{1}} \textrm{ln}sizemax +  \beta_{\textrm{2}} sst + \epsilon,$$
<!-- \label{kmax} -->
<!-- \end{equation} -->

\noindent where $\textrm{ln}kmax$ represents the natural log-transformed kmax value, $\beta_\textrm{0}$ is the fixed-effect intercept, $\gamma_\textrm{0phy}$ is the vector of random-effect coefficients that account for the residual intercept variation, based on the relatedness as described by the phylogeny, $\beta_\textrm{1}$ is the slope for the natural transformed maximum body size, $\beta_\textrm{2}$ is the slope for the average ambient sea surface temperature, $\epsilon$ is the residual variation.
\noindent We used uninformative priors and ran the model for 2000 iterations with a warm-up of 1000 iteration for 4 chains. 
\noindent The model fit confirmed a negative relationship of $\textrm{ln}kmax$ with $\textrm{ln}size$, and a positive relationship with sea surface temperature. The Bayesian R2 of the model was 0.738 (95\%CI: 0.702-0.769). 
\noindent The phylogenetic heritability (equivalent to Pagel’$\lambda$) was estimated as the proportion of total variance, conditioned on the effects, attributable to the phylogeny(i.e. $\lambda = \frac{{\textrm{sd}(\gamma_{\textrm{0phy}})}^2} {{\textrm{sd}(\gamma_\textrm{0phy})}^2 + \epsilon^2}$). This calculation resulted in a phylogenetic signal of 0.74 (95\% CI: 0.70 - 0.77).

\noindent We extrapolated $kmax$ for all species across the full temperature range in which those species occur in the database, with temperature rounded to the °C, which results in 4712 unique temperature and species combinations.  
There is currently no streamlined method to make predictions for new species from a phylogenetic regression model. We circumvented the issue by extracting draws of the phylogenetic effect, $\gamma_\textrm{0phy}$ for each species included in the model. We subsequently predicted these phylogenetic effects for missing species with the help of the function phyEstimate in the picante package for R [@Kembel2010]. This function uses phylogenetic ancestral state estimation to infer trait values for new species on a phylogenetic tree by rerooting the tree to the parent edge for the node to be predicted [@Kembel2012]. We repeated this for all 100 trees and 1000 draws. Per draw, we averaged the extrapolated values per species for the hundred trees. Then, by combining the predicted phylogenetic effects with the global intercept and slopes for body size and temperatures for each draw, we predicted $\kappa$ for each species. We only use one chain in order to keep computational time reasonable. Finally, we summarised all $\kappa$ predictions per sst per species by taking the mean and standard deviation across the 1000 draws.

## 2 Body stoichiometry    
### 2.1 Data collection    
\noindent 1633 individuals of 108 species and 25 families were collected between 2015 and 2017 in Mo'orea (Parravicini, unpub.), the Caribbean [@Allgeier2020sci], and Palmyra (Bukepile unpub.). Their gut contents were removed and the whole body was freeze-dried and ground to powder with a Precellys homogenizer. Whole body elemental proportions ($Q_\textrm{k}$) were then measured in the lab using standard methods. Specifically, ground samples were analyzed for \%C and \%N content using a CHN Carlo-Erba elemental analyzer (NA1500), and for \%P using dry oxidation-acid hydrolysis extraction followed by a colorimetric analysis [@Allen1974]. Elemental content was calculated based on dry mass.

### 2.2 Data analysis and extrapolation

\noindent The CNP\% content of organisms is known to be highly conserved within families [@Allgeier2020sci]. We therefore use phylogeny to extrapolate these values. We fitted C, N and P contents (\%) through a hierarchical phylogenetic multivariate normal model with phylogenetic effects and random effects per species.

<!-- \begin{equation} -->
$$\begin{bmatrix}Y_1 \\ Y_2 \\ Y_3\end{bmatrix} \sim MVNormal(\begin{bmatrix}mu1 \\ mu2 \\ mu3\end{bmatrix}, S),$$
<!-- 	\label{cnp} -->
<!-- \end{equation} -->

$$mu_{\textrm{n}\times\textrm{k}} = \beta_\textrm{0 k} + \gamma_{\textrm{0phy}\times\textrm{k}} + \gamma_{\textrm{0sp}\times\textrm{k}},$$
<!-- 	\label{cnp2} -->
<!-- \end{equation} -->

\noindent where $Y_1$, $Y_2$ and $Y_3$ are the \% content of $\textrm{C}$, $\textrm{N}$, and $\textrm{P}$ respectively, $mu_{\textrm{n}\times\textrm{k}}$ represents the average \% content of element $k$ ($\textrm{C}$, $\textrm{N}$, and $\textrm{P}$) per species, $\beta_\textrm{0 k}$ is the fixed-effect intercept for each element $k$, $\gamma_{\textrm{0phy}\times\textrm{k}}$ is the matrix of random-effect coefficients that account for the intercept variation, based on the relatedness as described by the phylogeny per element k, $\gamma_{\textrm{0sp}\times\textrm{k}}$ is the matrix of random-effect coefficients that account for the residual species-level intercept variation per element k.

\noindent We used uninformative priors and ran the model for 2000 iterations with a warm-up of 1000 iteration for 4 chains. 
\noindent The Bayesian R2 of the model was 0.39 (95\%CI: 0.36-0.42), 0.50 (95\%CI: 0.48-0.53), and 0.43 (95\%CI: 0.40-0.46) for C, N and P respectively. 
\noindent The phylogenetic heritability was 0.41 (95\%CI: 0.28-0.55), 0.58 (95\%CI: 0.4-0.66), and 0.57 (95\%CI: 0.46-0.69) for C, N, and P respectively.

\noindent As before, we used 1000 fitted draws for each species, and 100 phylogenetic trees to extrapolate to all species with unknown body stoichiometry. Specifically, we used the phylopars function from the Rphylopars package [@Bruggeman2009]. This function uses ancectral state reconstruction and brownian motion, and takes the correlation between C, N and P into account.

## 3 Diet     
### 3.1 Data collection 
\noindent We collected 571 adult individuals of 51 species between 2018 and 2019 in Mo'orea and Tetiaroa, and Mangareva, three Polynesian islands. We extracted the stomach content and stored it in a 2ml tube. After freezing the samples, we dry-froze all samples for at least 24 hours, and ground to powder. Then, samples were sent to the lab for CNP content analysis using similar methods as for the fish body stoichiometry described above.

### 3.2 Data analysis and extrapolation  
\noindent Using trophic guilds defined by @Parravicini2020, we fitted a multivariate Bayesian regression model to summarize CNP\% content data per trophic guild with random effects at the species level. This model had a median Bayesian R2 of 0.62, 0.62, and 0.48 for C, N and P respectively.           
Next, we extracted 1000 draws of the predicted CNP\% per trophic guild. Parravicini et al. 2020 [@Parravicini2020] provides the probability of reef fish species to be assigned to each of the eight defined trophic guilds(i.e. sessile invertivores; herbivores, microvores, and detrivores; corallivores; piscivores; microinvertivores; macroinvertivores; crustacivores; planktivores). By combining these probabilities with the predicted diet contents per trophic guild, we finally estimated the diet CNP\% for each species in our database. We then took the average and standard deviation across all 1000 draws. While we recognize the bias of using diet CNP\% estimates of a dataset in one region, we argue that variability between food categories e.g. animal material and primary producers is likely to be higher than regional differences within trophic categorizations. Further, since the used trophic guild classification includes probabilities to belong to each group, variation is included when the trophic categorization is not well known. For example, if a species has a 50% probability to be a herbivore and a 50% probability to be a sessile invertivore this uncertainty will be reflected the estimation of the diet CNP\%.

## 4 Metabolic parameters    
### 4.1 Data collection and lab experiments    
To quantify standard metabolic rate (SMR) and maximum metabolic rate (MMR), we carried out intermittent-closed respirometry experiments [@Steffensen1989; @Clark2013].
In the period between 2018 and 2019, we collected 1393 individuals of 61 species and 18 families with a minimum of 3 replicates per species using hand nets and clove oil (Parravicini, Brandl, and Mercière, unpub.).
After an acclimatization and fasting period of 48 h in aquaria, the fishes were individually transferred to a water-filled tub at 28°C (ambient sea water temperature) and manually chased by the experimenter until exhausted [@Norin2011; @Clark2012]. Then, they were placed in respirometry chambers submersed in an ambient and temperature-controlled tank, where they were left for ~23 h. The intermittent respirometry cycles started immediately after a fish was placed in its respirometry chamber. The cycles consisted of a measurement (sealed) period followed by a flush period during which the respirometry chambers were flushed with fully aerated water from the ambient tank. Because fishes were exhausted right before entering the respirometry chambers, it was possible to measure the approximate MMR. Depending on fish size, 8 respirometry chambers ranging in volume (including tubes and pumps) from 0.4 to 4.4 L were run in parallel, and measurement and flush periods lasted between 3 to 15 min and 3 to 5 min, respectively. SMR was calculated as the average of the 10 \% lowest values measured during the entire period, after the removal of outliers [@Chabot2016]. MMR was calculated from the slope of the first measurement period. \par

### 4.3 Data analysis and extrapolation
To retrieve the parameters $f_\textrm{0}$ (Metabolic normalisation constant independent of body mass; $\textrm{g C} \textrm{g}^{-\alpha} \textrm{d}^{-1}$), $\alpha$ (mass-scaling exponent), and $\theta$ (factorial activity scope), we fitted a normal Bayesian mixed effect model predicting the log10-transformed metabolic rate using the log10-transformed biomass including random effects of family, species, and mr type (SMR or MMR) on both the species and family intercept effect using brms [@Burkner2017]. We ran the model for 4000 iterations, with a warm-up of 2000 iterations, and four chains. Further, we used an informative prior for the slope ($\alpha \sim normal(0.8, 0.5)$. The model had a Bayesian R2 of 0.973 (95\%CI: 0.972-0.974).
We then predicted the family-level $\alpha$ by summing the slope of the model with the effects of the family on the slope of the SMR. We did this for all iterations and then took the mean and standard deviation. In a similar way, we extracted the family-level intercept for SMR and MMR (figure \@ref(fig:c4f4)), and then quantified mean and standard deviation of $f_\textrm{0}$ after the back-transformation of all iterations of the intercept (i.e., the intercept for SMR).
Finally, we estimated $\theta$ as followed, based on the assumption that fishes rest 12h a day and they on average spend the remaining 12 hours at a metabolic rate that is the average of their SMR and MMR:
<!-- \begin{equation} -->
$$\theta = \frac{3\textrm{SMR} + \textrm{MMR}}{4\textrm{SMR}},$$
<!-- 	\label{theta} -->
<!-- \end{equation} -->
where all iterations of the back-transformed family-level intercepts were used for SMR and MMR. We then summarized these predictions by taking the mean and standard deviation.
We used the family-level estimates for these parameters for all species in our database. For families that were not represented in our respirometry dataset, we used the average across all families. The parameter $f0$ was adjusted for each temperature in the varying locations of the underwater visual census database, following @Barneche2014:

$$f0_{adjusted} = f0(T_{ref}) e^{E_r(\frac{1}{\kappa T_ref} - \frac{1}{\kappa T})},$$
where $\kappa$ is the Boltsmann's constant ($8.62 * 10^{-5} eV K^{-1}$), $E_r$ is the activation energy (0.59 eV, extracted from @Barneche2014), $f0(T_{ref})$ is the value of the metabolic normalization at a fixed reference temperature (301.15 K = 28 + 273.15K), $T$ is the temperature of interest (K).

## 5. Additional parameters  
We retrieved the parameters $lw_a$, $lw_b$, $h$, and $r$ from fishbase [@Froese2018]. For the mass-specific turnover rates for N and P($F_\textrm{0Nz}$; $F_\textrm{0Pz}$), we used the estimates provided in Schiettekatte et al. (2020) [@Schiettekatte2020].

## References