### 2021-10-06 text/suppl_tables_figures.Rmd
:fish:

---
output:
  word_document:
    fig_caption: yes
    reference_docx: ../text/word_template.docx
bibliography: ../text/global_proc.bib
csl: nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE)
library(tidyverse)
library(flextable)
library(officer)
library(drake)
library(broom)

# loadd(SI_fig1)
# loadd(SI_fig2)
# loadd(SI_pp_plots)
# loadd(SI_rank_plots)
```

# Supplementary data figures and tables: Drivers and vulnerability of global coral reef fish functions

## Tables

```{r, echo = FALSE}
data <- drake::readd(summary_transect_complete) %>%
  group_by(bioregion, locality) %>%
  summarize(n_sites = length(unique(sites)),
            n_transects = n())

  flextable(data) %>%
    fontsize(size = 8) %>%
    fontsize(size = 10, part = "header") %>%
    set_caption("Extended Date table 1: Overview of localities of UVC transects, used in this study, including number of sites and number of transects") %>%
    autofit()
```

```{r, echo = FALSE}
 table <- readd(tab_mod_mv_siteloc)
  flextable(table) %>%
    fontsize(size = 8) %>%
    fontsize(size = 10, part = "header") %>%
    merge_v(j = 1) %>% 
    border_inner_h(border = fp_border(color = "black", style = "solid", width = 1)) %>%
    set_caption("Supplementary table 2: Overview of parameters of the multivariate regression predicting the five functions with intercepts only.") %>%
    autofit()
```

```{r, echo = FALSE}
 table <- readd(tab_mod_mvfun_bm)
  flextable(table) %>%
    fontsize(size = 8) %>%
    fontsize(size = 10, part = "header") %>%
    merge_v(j = 1) %>% 
    border_inner_h(border = fp_border(color = "black", style = "solid", width = 1)) %>%
    set_caption("Supplementary table 3: Overview of parameters of the multivariate regression predicting the five functions with biomass and sst.") %>%
    autofit()
```

```{r, echo = FALSE}
 table <- readd(tab_mod_mvfun_com)
  flextable(table) %>%
    fontsize(size = 8) %>%
    fontsize(size = 10, part = "header") %>%
    merge_v(j = 1) %>% 
    border_inner_h(border = fp_border(color = "black", style = "solid", width = 1)) %>%
    set_caption("Supplementary table 4: Overview of parameters of the multivariate regression predicting the five functions and community variables.") %>%
    autofit()
```

## Figures

![Extended Data Figure 1: Correlations among functions, independent of biomass and sst, on the locality and site level.](../output/plots/figSx_corplot.png)

![Extended Data Figure 2: Posterior predictive checks of the five models relating functions  with biomass and sea surface temperature only (a) N excretion, b) P excretion, c) Production, d) Herbivory, e) Piscivory), and the five models relating functions with community variables (f) N excretion, g) P excretion, h) Production, i) Herbivory, j) Piscivory)](../output/plots/pp_plots.png)

![Extended Data Figure 3: Average relative contribution of fish families to all five functions per biogeographical ocean basin. CIP = Central-Indo-Pacific, CP = Central Pacific, EA = Eastern Atlantic, WA = Western Atlantic, WI = Western Indian](../output/plots/con_family_rank.png)
![Extended Data Figure 4: Average relative contribution of the top ten most contributing species to all five functions per biogeographical ocean basin. CIP = Central-Indo-Pacific, CP = Central Pacific, EA = Eastern Atlantic, WA = Western Atlantic, WI = Western Indian](../output/plots/con_species_rank.png)

![Extended Data Figure 5: Comparison herbivory and piscivory rates when using alternative diet classifications from Mouillot et al. (2014) and Siqueira et al. 2020](../output/plots/Supplemental_herbivory_piscivory_alternative_classification.png)